---
title: "Single-cell RNAseq analysis of gastruloids"
subtitle: "urd | velocity"
author:
- Thomas Vannier
date: "`r format(Sys.time(), '%d %B %Y')`"
output:
    html_document:
        theme:
            bootswatch: yeti
        toc: yes
        toc_float:
          collapsed: false
          smooth_scroll: true
        number_sections: yes
        df_print: kable
        code_folding: hide
---

<!-- Javascript for zooming on figures (adapted from: https://stackoverflow.com/questions/40401680) -->

<!-- Jquery import conflicts with DT::datatable so needs to be commented here -->
<!-- <script src="https://ajax.googleapis.com/ajax/libs/jquery/3.1.1/jquery.min.js"></script> -->

<style>
.zoomDiv {
  display: none;
  position: fixed;
  top: 50%;
  left: 50%;
  z-index: 50;
  transform: translate(-50%, -50%);
  background-color: #FFFFFF;
  box-shadow: 0px 0px 50px #888888;
  width: fit-content;
  max-width: 90%;
  max-height: 90%;
  overflow: auto;
}

.zoomImg {
  width: 150%;
}
</style>

<script type="text/javascript">
  $(document).ready(function() {
    $('body').prepend("<div class=\"zoomDiv\"><img src=\"\" class=\"zoomImg\"></div>");
    // onClick for all img except the zoomed one and link ones (filter)
    // use 'img.zoom' and out.extra='class=\"zoom\"' in chunk to specify manually which chunk images can be zoomed
    $('img:not(.zoomImg)').filter(':not(a *)').click(function() {
      $('.zoomImg').attr('src', $(this).attr('src'));
      $('.zoomDiv').show();
    })
    // onClick function for hiding div
    $('img.zoomImg').click(function() {
      $('.zoomDiv').hide();
    })
  })
</script>

# This report present the velocity analysis with the urd tools.

URD : URD is an R package for reconstructing transcriptional trajectories underlying 
specification or differentiation processes in the form of a branching tree,
using single cell RNA-sequencing data. URD uses a diffusion map projection
of the data and works by simulating biased random walks through that projection. 

I follow the quick start guide from URD github page :
(https://github.com/farrellja/URD/blob/master/Analyses/QuickStart/URD-QuickStart-AxialMesoderm.md)
and the workflow of Céline Chevalier :
https://github.com/BAUDOTlab/gastruloid_timeserie_scRNA-seq

```{r setup, warnings=F}
knitr::opts_chunk$set(echo = TRUE, error = TRUE)
```

```{r loadLibraries, echo=FALSE, warning=FALSE, message=FALSE}
# Installing packages when not possible with conda env

options(repos = c(CRAN = "https://cloud.r-project.org/"))
install.packages("BiocManager")
BiocManager::install("GenomeInfoDbData")
devtools::install_github("farrellja/URD")

# Loading library

suppressPackageStartupMessages(library(rgl))
suppressPackageStartupMessages(library(URD))
suppressPackageStartupMessages(library(ggplot2))
suppressPackageStartupMessages(library(kableExtra))
library(rmarkdown)
library(tidyverse)
library(stringr)
library(tinytex)
library(knitr)
```

# Load the data

For this analysis, I use the merged data of the 4 times (ie. merged_urd_object.rds). I get an urd-object by running merge_rds.R.
```{r loadData}
# Load the 4 times merged dataset
TEXT_OUTPUT = snakemake@output[["urd_output"]]
URD_OBJECT = snakemake@params[["urd_object"]]
object <- readRDS(URD_OBJECT)
# str(object)
counts_matrix <- object@count.data
counts_matrix[1, 1:5] # For raw counts
logupx_matrix <- object@logupx.data
logupx_matrix[1, 1:5]
```

# Calculate variable features

CV: coefficient of variation

```{r CalcVariableFeatures}
stages <- sort(unique(object@group.ids$SampleID))

cells.each.stage <- lapply(stages,
                           function(stage){
                               rownames(object@group.ids)[which(object@group.ids$SampleID == stage)]
                           }
)

var.genes.by.stage <- lapply(1:length(stages),
                             function(n){
                                 findVariableGenes(object,
                                                   cells.fit = cells.each.stage[[n]],
                                                   set.object.var.genes = F,
                                                   diffCV.cutoff = 0.1,
                                                   mean.min = 0.005,
                                                   mean.max = 100,
                                                   main.use = stages[[n]],
                                                   do.plot = T
                                 )
                             }
)

# combine variable features
names(var.genes.by.stage) <- stages
var.genes <- sort(unique(unlist(var.genes.by.stage)))
object@var.genes <- var.genes
```

```{r CalcPCA}
seed <- snakemake@params[["seed"]]
set.seed(seed)
# Calculate PCA and consider those PCs that with standard deviation 'seed'x expected by noise as significant
object <- calcPCA(object, mp.factor = seed)
pcSDPlot(object)
```

```{r CalcTsne}
# Remove cells with duplicates PCA coordinates
cellsToRemove <- rownames(object@pca.scores[duplicated(object@pca.scores),])
print(cellsToRemove)
object <- urdSubset(object,
                    cells.keep = setdiff(colnames(object@logupx.data), cellsToRemove)
)

# Perform Tsne projection
set.seed(seed)
object <- calcTsne(object,
                   perplexity = 30,
                   theta=0.5
)

plotDim(object, "SampleID", plot.title = "tSNE: Time")
```

# Remove outliers

```{r Output}
# Create the output file for the snakemake rule

TEXT_OUTPUT <- snakemake@output[["urd_output"]]

output_file<-file(TEXT_OUTPUT)
writeLines(c("Rules URD finished"), output_file)
close(output_file)
```
