---
title: "Single-cell RNAseq analysis of the somatosensory cortex"
subtitle: "Seurat | QC and Clustering"
author:
- Léa Chabot
date: "`r format(Sys.time(), '%d %B %Y')`"
output:
    html_document:
        theme: cerulean
        toc: true
        toc_float: true
        self_contained: true
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE, error=TRUE)
```

```{r loadLibraries, echo=FALSE, warning=FALSE, message=FALSE}
library(rmarkdown)
library(tidyverse)
library(stringr)
library(tinytex)
library(knitr)
library(readr)
library(dplyr)
library(Seurat)
library(patchwork)
library(sctransform)
library(glmGamPoi)
library(ggplot2)
library(data.table)
library(devtools)
library(sessioninfo)
library(dittoSeq)
library(DT)
library(ggpubr)
library( ggrepel)
library( RColorBrewer)
library( scales)
library(plotly)
library( htmltools)
```

[//]: # "Load the various chunks in the various files"
```{r readChunks, echo=FALSE, warning=FALSE, message=FALSE}
read_chunk( path=file.path( SCRIPTDIR, "01_prepare_data.R"));
read_chunk( path=file.path( SCRIPTDIR, "02_variables_genes.R"));
read_chunk( path=file.path( SCRIPTDIR, "03_cell_heterogeneity_PCA.R"));
read_chunk( path=file.path( SCRIPTDIR, "04_cell_heterogeneity_tSNE_UMAP.R"));
read_chunk( path=file.path( SCRIPTDIR, "05_cell_heterogeneity_clustering.R"));
read_chunk( path=file.path( SCRIPTDIR, "06_markers.R"));
```

This report present the analyse of dataset generated with 10x Genomics Single Cell Protocol.

# Primary analysis : QC and selecting cells for further analysis of the sample `r SAMPLE_ID`

```{r loadData, echo = FALSE, fig.height = 10, fig.width = 10, out.height = 600, out.width = 600, results = "asis", message = FALSE}
<<loadData>>
```
## Filtering cells

We visualize QC metrics, and use these to filter cells.

- We filter cells that have unique UMI counts over `r FILTER_UMI_MAX` (different number corresponding to the different samples) or less than `r FILTER_UMI_MIN`

- We filter cells that have unique feature counts over `r FILTER_FEATURE_MAX` (different number corresponding to the different samples) or less than `r FILTER_FEATURE_MIN`

- We filter cells that have > `r FILTER_PERCENT_MT`% mitochondrial counts or or less than `r FILTER_PERCENT_MT_MIN`.

- We filter cells that have > `r FILTER_PERCENT_RB`% ribosomal counts.


```{r filterData_selection, echo = FALSE, fig.height = 10, fig.width = 10, out.height = 600, out.width = 600, results = "asis", message = FALSE}
<<filterData_selection>>
```

```{r filterData_summaryPlot, echo = FALSE, fig.height = 10, fig.width = 10, out.height = 600, out.width = 600, results = "asis", message = FALSE}
<<filterData_summaryPlot>>
```

```{r filterData_filterObject, echo = FALSE, fig.height = 10, fig.width = 10, out.height = 600, out.width = 600, results = "asis", message = FALSE}
<<filterData_filterObject>>
```

## Normalizing the data

After removing unwanted cells from the dataset we normalize the feature expression measurements for each cell (and correct the batch effect using the Seurat integration procedure if necessary) with `r snakemake@params[["norm_method"]]`
We apply a linear transformation (‘scaling’) that is a standard pre-processing step prior to dimensional reduction techniques like PCA. If SCTransform is performed, it's not necessary to scale the data after it.

```{r normalizeData, echo = FALSE, fig.height = 10, fig.width = 10, out.height = 600, out.width = 600, results = "asis", message = FALSE}
<<normalizeData>>
```

<br /><br />

## Identification of highly variable genes

Subset of features that exhibit high cell-to-cell variation in the dataset

```{r findVariableGenes_seuratMethod, echo = FALSE, fig.height = 10, fig.width = 10, out.height = 600, out.width = 600, results = "asis", message = FALSE, warning = FALSE}
<<findVariableGenes_seuratMethod>>
```

Number of variable genes found: `r variablesGenesStats`

```{r findVariableGenes_summaryPlot, echo = FALSE, fig.height = 10, fig.width = 10, out.height = 600, out.width = 600, results = "asis", message = FALSE, warning = FALSE}
<<findVariableGenes_summaryPlot>>
```

<br />

Top `r VARIABLE_FEATURES_SHOWTOP` most variable annotations:

```{r findVariableGenes_summaryTable, echo = FALSE, fig.height = 10, fig.width = 10, out.height = 600, out.width = 600, results = "asis", message = FALSE, warning = FALSE}
<<findVariableGenes_summaryTable>>
```

```{r scaleData, echo = FALSE, fig.height = 10, fig.width = 10, out.height = 600, out.width = 600, results = "asis", message = FALSE}
<<scaleData>>
```

<br /><br />

```{r, echo = FALSE}
if (as.logical(MULTIPLEX) == TRUE) {
  include_chunk <- TRUE
  include_title <- TRUE
} else {
  include_chunk <- FALSE
  include_title <- FALSE
}
```

`r if(include_title) "## Demultiplexing"`
```{r HTODemux, echo = FALSE, fig.height = 10, fig.width = 10, out.height = 600, out.width = 600, results = "asis", message = FALSE}
if (include_chunk) {<<HTODemux>>}
```

<br /><br />

## PCA results

Visualizing both cells and features that define the PCA

```{r ldr, echo=FALSE,fig.height=20,fig.width=20,out.height=600,out.width=600, results = "asis", message=FALSE}
<<ldr>>
```

## Determine the ‘dimensionality’ of the dataset

How many components have to be include?

### Elbow Plot

A ranking of principle components based on the percentage of variance explained by each one.

```{r elbowplot, echo=FALSE, message=FALSE}
<<elbowplot>>
```

Number of dimensions to use for PCA : `r snakemake@params[["pca_npc"]]``

PCA projections for dimensions `r paste( 1:PCA_PLOTS_NBDIMS, collapse="/")` (colored by batch)  

```{r heterogeneity_pca, echo = FALSE, out.width='33%', results = "asis", message = FALSE, warning = FALSE}
<<heterogeneity_pca>>
```

<br />

UMIs counts on PCA dimensions `r paste( 1:PCA_PLOTS_NBDIMS, collapse="/")`  
```{r heterogeneity_pca_umisCounts, echo = FALSE, out.width='33%', results = "asis", message = FALSE, warning = FALSE}
<<heterogeneity_pca_umisCounts>>
```

<br />

Genes counts on PCA dimensions `r paste( 1:PCA_PLOTS_NBDIMS, collapse="/")`     
```{r heterogeneity_pca_genesCounts, echo = FALSE, out.width='33%', results = "asis", message = FALSE, warning = FALSE}
<<heterogeneity_pca_genesCounts>>
```

<br />

Correlation (Spearman) of PCA dimensions `r paste( 1:PCA_PLOTS_NBDIMS, collapse="/")` with UMIs and genes counts  
```{r heterogeneity_pca_correlations, echo = FALSE, fig.dim = c(9,6), results = "asis", message = FALSE, warning = FALSE}
<<heterogeneity_pca_correlations>>
```
<br />

PCA loadings of top `r PCA_PLOTS_NBFEATURES` features for dimensions `r paste( 1:PCA_PLOTS_NBDIMS, collapse="/")`  
```{r heterogeneity_pca_loadings, echo = FALSE, fig.dim = c(4, 4), out.width='33%', results = "asis", message = FALSE, warning = FALSE}
<<heterogeneity_pca_loadings>>
```
<br /><br />

## Dimensionality reduction

```{r heterogeneity_dimReduc, echo = FALSE, fig.height = 10, fig.width = 10, out.height = 600, out.width = 600, results = "asis", message = FALSE, warning = FALSE}
<<heterogeneity_dimReduc>>
```

### Colored by batch {.tabset .tabset-fade}

#### UMAP
```{r dimreduc_ggplot_covariables, echo = FALSE, fig.height = 10, fig.width = 10, out.height = 600, out.width = 600, results = "asis", message = FALSE}
useReduction='umap'
<<dimreduc_ggplot_covariables>>
rm("useReduction");
```

#### TSNE
```{r dimreduc_ggplot_covariables_tSNE, echo = FALSE, fig.height = 10, fig.width = 10, out.height = 600, out.width = 600, results = "asis", message = FALSE}
useReduction='tsne'
<<dimreduc_ggplot_covariables_tSNE>>
rm("useReduction");
```

<br /><br />


## Clusters identification

We use a graph-based clustering approach : 1. Construct a KNN graph based on the euclidian distance in PCA space, and refine the edge weights between any two cells based on the shared overlap in their local neighborhoods (Jaccard similarity). 2. Modularity optimization techniques such as the Louvain algorithm or SLM, to iteratively group cells together, with the goal of optimizing the standard modularity function.

### Cells distribution
```{r heterogeneity_identifyClusters, echo = FALSE, fig.height = 10, fig.width = 10, out.height = 600, out.width = 600, results = "asis", message = FALSE}
<<heterogeneity_identifyClusters>>
```

### Statistics by cluster
```{r heterogeneity_identifyClusters_splitStats, echo = FALSE, fig.height = 10, fig.width = 10, out.height = 600, out.width = 600, results = "asis", message = FALSE, warning = FALSE}
<<heterogeneity_identifyClusters_splitStats>>
```
<br /><br />

### Clusters projection {.tabset .tabset-fade}

#### UMAP
```{r heterogeneity_dimReduc_with_clusters, echo = FALSE, fig.height = 10, fig.width = 10, out.height = 600, out.width = 600, results = "asis", message = FALSE}
useReduction='umap'
<<heterogeneity_dimReduc_with_clusters>>
rm("useReduction");
```

#### tSNE
```{r heterogeneity_dimReduc_with_clusters_tSNE, echo = FALSE, fig.height = 10, fig.width = 10, out.height = 600, out.width = 600, results = "asis", message = FALSE}
useReduction='tsne'
<<heterogeneity_dimReduc_with_clusters_tSNE>>
rm("useReduction");
```

### {-}

<br /><br />

```{r save_RDS, echo = FALSE, fig.height = 10, fig.width = 10, out.height = 600, out.width = 600, results = "asis", message = FALSE, warning = FALSE}
<<save_RDS>>
```


##  GABAergic markers {.tabset .tabset-fade}
Here you can vizualize GABAergic markers expression.

### GABAN
```{r GABAN, echo = FALSE, fig.height = 10, fig.width = 10, out.height = 600, out.width = 600, results = "asis", message = FALSE}
<<GABAN>>
```

### Meis2
```{r Meis2, echo = FALSE, fig.height = 10, fig.width = 10, out.height = 600, out.width = 600, results = "asis", message = FALSE}
<<Meis2>>
```

### Lamp A7C|CNC
```{r Lamp5_A7C_CNC, echo = FALSE, fig.height = 10, fig.width = 10, out.height = 600, out.width = 600, results = "asis", message = FALSE}
<<Lamp5_A7C_CNC>>
```

### Lamp A7C|NGC
```{r Lamp5_A7C_NGC, echo = FALSE, fig.height = 10, fig.width = 10, out.height = 600, out.width = 600, results = "asis", message = FALSE}
<<Lamp5_A7C_NGC>>
```

### Sngc CCK|BC
```{r Sncg_CCK_BC, echo = FALSE, fig.height = 10, fig.width = 10, out.height = 600, out.width = 600, results = "asis", message = FALSE}
<<Sncg_CCK_BC>>
```

### Vip BP|BTC
```{r Vip_BP_BTC, echo = FALSE, fig.height = 10, fig.width = 10, out.height = 600, out.width = 600, results = "asis", message = FALSE}
<<Vip_BP_BTC>>
```

### Sst T-MC
```{r Sst_T_MC, echo = FALSE, fig.height = 10, fig.width = 10, out.height = 600, out.width = 600, results = "asis", message = FALSE}
<<Sst_T_MC>>
```

### Sst fan-MC
```{r Sst_fan_MC, echo = FALSE, fig.height = 10, fig.width = 10, out.height = 600, out.width = 600, results = "asis", message = FALSE}
<<Sst_fan_MC>>
```

### Sst NMC
```{r Sst_NMC, echo = FALSE, fig.height = 10, fig.width = 10, out.height = 600, out.width = 600, results = "asis", message = FALSE}
<<Sst_NMC>>
```

### Sst IVC
```{r Sst_IVC, echo = FALSE, fig.height = 10, fig.width = 10, out.height = 600, out.width = 600, results = "asis", message = FALSE}
<<Sst_IVC>>
```

### Sst FS-like
```{r Sst_FS_like, echo = FALSE, fig.height = 10, fig.width = 10, out.height = 600, out.width = 600, results = "asis", message = FALSE}
<<Sst_FS_like>>
```

### Sst Chold|LPC
```{r Sst_Chold_LPC, echo = FALSE, fig.height = 10, fig.width = 10, out.height = 600, out.width = 600, results = "asis", message = FALSE}
<<Sst_Chold_LPC>>
```

### Pvalb CHC
```{r Pvalb_CHC, echo = FALSE, fig.height = 10, fig.width = 10, out.height = 600, out.width = 600, results = "asis", message = FALSE}
<<Pvalb_CHC>>
```

### Pvalb FS|BC
```{r Pvalb_FS_BC, echo = FALSE, fig.height = 10, fig.width = 10, out.height = 600, out.width = 600, results = "asis", message = FALSE}
<<Pvalb_FS_BC>>
```

## {-}

##  Glutamatergic markers {.tabset .tabset-fade}
Here you can vizualize Glutamatergic markers expression.

### Cajal retzius cells
```{r Cajal_Retzius_cells_markers, echo = FALSE, fig.height = 10, fig.width = 10, out.height = 600, out.width = 600, results = "asis", message = FALSE}
<<Cajal_Retzius_cells_markers>>
```

### L2/3 IT
```{r L2_3_IT, echo = FALSE, fig.height = 10, fig.width = 10, out.height = 600, out.width = 600, results = "asis", message = FALSE}
<<L2_3_IT>>
```

### L4 IT SSC
```{r L4_IT_SSC, echo = FALSE, fig.height = 10, fig.width = 10, out.height = 600, out.width = 600, results = "asis", message = FALSE}
<<L4_IT_SSC>>
```

### L4 IT PC
```{r L4_IT_PC, echo = FALSE, fig.height = 10, fig.width = 10, out.height = 600, out.width = 600, results = "asis", message = FALSE}
<<L4_IT_PC>>
```

### L5 IT
```{r L5_IT, echo = FALSE, fig.height = 10, fig.width = 10, out.height = 600, out.width = 600, results = "asis", message = FALSE}
<<L5_IT>>
```

### L5 PT
```{r L5_PT, echo = FALSE, fig.height = 10, fig.width = 10, out.height = 600, out.width = 600, results = "asis", message = FALSE}
<<L5_PT>>
```

### L5/6 NP
```{r L5_6_NP, echo = FALSE, fig.height = 10, fig.width = 10, out.height = 600, out.width = 600, results = "asis", message = FALSE}
<<L5_6_NP>>
```

### L6 IT
```{r L6_IT, echo = FALSE, fig.height = 10, fig.width = 10, out.height = 600, out.width = 600, results = "asis", message = FALSE}
<<L6_IT>>
```

### L6 CT
```{r L6_CT, echo = FALSE, fig.height = 10, fig.width = 10, out.height = 600, out.width = 600, results = "asis", message = FALSE}
<<L6_CT>>
```

### L6b
```{r L6b, echo = FALSE, fig.height = 10, fig.width = 10, out.height = 600, out.width = 600, results = "asis", message = FALSE}
<<L6b>>
```

## {-}

##  Non-Neuronal markers {.tabset .tabset-fade}
Here you can vizualize GABAergic markers expression.

### Astrocytes
```{r astro_markers, echo = FALSE, fig.height = 10, fig.width = 10, out.height = 600, out.width = 600, results = "asis", message = FALSE}
<<astro_markers>>
```

### Oligodendrocytes
```{r oligo_markers, echo = FALSE, fig.height = 10, fig.width = 10, out.height = 600, out.width = 600, results = "asis", message = FALSE}
<<oligo_markers>>
```

### Microglia
```{r microglia_markers, echo = FALSE, fig.height = 10, fig.width = 10, out.height = 600, out.width = 600, results = "asis", message = FALSE}
<<microglia_markers>>
```

### Endothelial
```{r Endothelial_cells_markers, echo = FALSE, fig.height = 10, fig.width = 10, out.height = 600, out.width = 600, results = "asis", message = FALSE}
<<Endothelial_cells_markers>>
```

### Pericyte
```{r Pericyte_markers, echo = FALSE, fig.height = 10, fig.width = 10, out.height = 600, out.width = 600, results = "asis", message = FALSE}
<<Pericyte_markers>>
```

## {-}
