---
title: "Single-cell RNAseq analysis of the human melanoma cells | IBDM | Marie-Claire Delfini"
subtitle: "Seurat | QC and Clustering"
author:
- Thomas Vannier | CENTURI Multi-Engineering Platform
date: "`r format(Sys.time(), '%d %B %Y')`"
output:
  html_document: 
    theme: cerulean
    toc: true
    toc_float: true
    self_contained: true
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE, error=TRUE)
```

```{r loadLibraries, echo=FALSE, warning=FALSE, message=FALSE}
library(rmarkdown)
library(tidyverse)
library(stringr)
library(tinytex)
library(knitr)
library(readr)
library(dplyr)
library(Seurat)
library(patchwork)
library(sctransform)
library(glmGamPoi)
library(ggplot2)
library(data.table)
library(devtools)
library(sessioninfo)
```

[//]: # "Load the various chunks in the various files"
```{r readChunks, echo=FALSE, warning=FALSE, message=FALSE}
read_chunk( path=file.path( SCRIPTDIR, "assignment.R"))
read_chunk( path=file.path( SCRIPTDIR, "seurat.R"))
```
This report reproduced the analyses from the [Guided Clustering Tutorial](https://satijalab.org/seurat/articles/pbmc3k_tutorial.html) of Seurat.

# QC and selecting cells for further analysis of the BioProject `r snakemake@params[["biop"]]`

## Visualize QC metrics as a violin plot and FeatureScatter

We visualize QC metrics, and use these to filter cells.

- We filter cells that have unique feature counts over `r snakemake@params[["maxFeature_RNA"]]`  or less than `r snakemake@params[["minFeature_RNA"]]`

- We filter cells that have >`r snakemake@params[["percent_mt"]]`% mitochondrial counts.

```{r qcmetrics, echo=FALSE,fig.height=10,fig.width=10,out.height=600,out.width=600, results = "asis", message=FALSE}
<<qcmetrics>>
```

## Normalizing the data

After removing unwanted cells from the dataset we normalizes the feature expression measurements for each cell using the `r snakemake@params[["normwf"]]` normalization workflow.

## Identification of highly variable features

Subset of features that exhibit high cell-to-cell variation in the dataset

```{r hvf, echo=FALSE,fig.height=10,fig.width=8,out.height=600,out.width=450, results = "asis", message=FALSE, warning=FALSE}
<<hvf>>
```

## Scaling the data

We apply a linear transformation (‘scaling’) that is a standard pre-processing step prior to dimensional reduction techniques like PCA.

## Perform linear dimensional reduction

Visualizing both cells and features that define the PCA

```{r ldr, echo=FALSE,fig.height=10,fig.width=10,out.height=600,out.width=600, results = "asis", message=FALSE}
<<ldr>>
```

## Determine the ‘dimensionality’ of the dataset

How many components have to be include?

### Elbow Plot

A ranking of principle components based on the percentage of variance explained by each one.

```{r elbowplot, echo=FALSE, message=FALSE}
<<elbowplot>>
```

## Cluster the cells

We use a graph-based clustering approach :
1. Construct a KNN graph based on the euclidean distance in PCA space, and refine the edge weights between any two cells based on the shared overlap in their local neighborhoods (Jaccard similarity).
2. Modularity optimization techniques such as the Louvain algorithm or SLM, to iteratively group cells together, with the goal of optimizing the standard modularity function. 

```{r cluster, echo=FALSE, fig.height=10, fig.width=10, out.height=600, out.width=600, results = "asis", message=FALSE}
<<cluster>>
```

## Run non-linear dimensional reduction

UMAP to visualize and explore these datasets.

```{r umap, echo=FALSE, fig.height=10, fig.width=10, out.height=600, out.width=600, results = "asis", message=FALSE}
<<umap>>
```

## Retrieve data in matrix

Count, expression and normalized matrice can be found in the snakemake report

```{r datasave, echo=FALSE, message=FALSE}
<<datasave>>
```

## Session Info

```{r info, echo=FALSE, message=FALSE}
<<info>>
```